Modal Abstractions for Virtualizing Memory Addresses

### 2024-08-05 examples.tex
halfway through updating figure 11

shr    0x27,rdi
and    0x1ff,rdi
shl    0x3,rdi
add    rdi,rsi
mov    rsi,-0x18[rbp]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry}\},m)  }_{\rtv}$
$\specline{ \ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{\begin{array}{l} \textsf{entry\_present(l4e\_val)} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4)}_{pfn} \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}    \end{array}}_{\rtv} $
$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{plm4 l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$    

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry}\},m)  }_{\rtv}$
$\specline{ \ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}}  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} } $
$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{plm4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})}_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry},\textsf{pdp\_entry}\},m)  }_{\rtv}$
$\specline{\ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
$\specline{ \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)

@@

\todo[inline]{colin frontier, need to incorporate these next 3 (2+1) paragraphs taken from earlier}
\todo[inline,color=green]{need to finish updating ghost token assertions in figure}

In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in S
pecification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}. Intuitively speaking, we start traversal from a certain entry address at the current level for which we know ($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$  Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.

@@

\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}). With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}. 
Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
of the value in memory --- so full ownership is returned only for unused entries.
This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.

@@

 $\specline{\textsf{P} \ast \textsf{plm4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      

@@

$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{plm4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
\ghostmaptoken{\textsf{id}}{(\mathsf{plm4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(plm4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{plm4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$

>>>>

shr    0x27,rdi       ;; Shift L4 index to lowest bits
and    0x1ff,rdi      ;; Mask to just lower 9 bits (0x1ff=511)
shl    0x3,rdi        ;; Multiply by 8
add    rdi,rsi        ;; Add to pml4 (virtual) table base
mov    rsi,-0x18[rbp] ;; Store to local variable pml4_entry; logical pml4_entry is physical, program variable is virtual
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m) \ast \forall_{i\in\textsf{0..511}\setminus\{\textsf{PML4EX(va)}\}}\ghostmaptoken{\textsf{id}}{(\textsf{rtv}+i*8)}{\textsf{4}}\ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{\textsf{entry\_present(l4e\_val)} \wand \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4)}_{pfn} \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv} $
$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$    

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry}\},m)  \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(l4e\_val'.pfn) + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \ast \ulcorner\textsf{pdp-KERNBASE}=\textsf{table\_root(l4e\_val'.pfn)}\urcorner} $
$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{pml4\_entry l4e\_val'} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner}_{\rtv}$
$\specline{\texttt{rax}  \mapsto_{\textsf{r}} \textsf{table\_root(l4e\_val'.pfn)} }_{\rtv}$ ;; pte_get_next_table may have allocated a new page, updating entry

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry}\},m)  }_{\rtv}$
$\specline{\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$
$\specline{\forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3} }_{\rtv}$
$\specline{ \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv}$(*\label{line:ex_l4_vpte}*)

@@

\todo[inline]{colin frontier, need to incorporate these next 3 (2+1) paragraphs taken from earlier}
\todo[inline,color=green]{need to finish updating ghost token assertions in figure; I think they're correct now up to the pdp_entry assignment line}

In the very first precondition on the client side -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir} -- 
we see that \textsf{vpte-ptsto} for the entry being accessed (\textsf{entry}) is constructed -- in 
Specification Line \ref{line:get_next_vpte_precondition} of Figure \ref{fig:calltopteinitialize} for \textsf{entry}. 
Intuitively speaking, we start traversal from a certain entry address at the current level for which we know 
($\textsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Line 3 in Figure \ref{walkpgdir}) that 
there is a memory mapping ($\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}$ 
Line 15 in Figure \ref{walkpgdir}) that can be used to explore deeper in the page-tree.

@@

\textsf{walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry})
 in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned
 token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation 
$\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained 
by offsetting the physical address (\textsf{entry}). With this knowledge on the root-page-table-entry, we can start traversing the page-table 
tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltopteinitialize}. 
Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a 
random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership 
returned must be determined by inspecting the valid bit
of the value in memory --- so full ownership is returned only for unused entries.
This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.

@@

 $\specline{\textsf{P} \ast \textsf{pml4} \mapsto_{\textsf{id}} \textsf{\_} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}}_{\rtv}$      

@@

$\specline{ \left( \begin{array}{l} \textsf{pdp}+\textsf{PDPEX(va)} = \textsf{pdp\_entry} \land \textsf{table\_root(pml4\_entry}_{\textsf{pfn}})  = \textsf{pdp} \\ \textsf{pd}+\textsf{PDEX(va)} = \textsf{pd\_entry} \land \textsf{table\_root(pml3\_entry}_{\textsf{pfn}})  = \textsf{pd} \\ \textsf{pt}+\textsf{PDPEX(va)} = \textsf{pt\_entry} \land \textsf{table\_root(pml2\_entry}_{\textsf{pfn}})  = \textsf{pt} \\  \textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus\{\textsf{pml4\_entry},\textsf{pdp\_entry},\textsf{pd\_entry},\textsf{pt\_entry}\},m)  \\ \ghostmaptoken{\delta{}s}{\rtv}{\delta} \\
\ghostmaptoken{\textsf{id}}{(\mathsf{pml4\_entry})}{\textsf{4}}\ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm3\_entry})}{\textsf{3}} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{plm2\_entry})}{\textsf{2}}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pml4\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{3}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pdp\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{2}  \\ \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pd\_entry}_{\textsf{pfn}}) \textsf{ + i * 8} \mapsto_{\textsf{id}} \textsf{1}  \\  \textsf{pml4\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l4e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l4e\_val})\urcorner \\  \textsf{pdp\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l3e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l3e\_val})\urcorner \\  \textsf{pd\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l2e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l2e\_val})\urcorner \\   \textsf{pt\_entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{l1e\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present l1e\_val})\urcorner    \end{array} \right)  =\textsf{R}_{\textsf{walk\_acc}}}$
